Decorrelating Errors in Quantum Gates by Random Gate Synthesis

### 2018-09-25 decorrelation/decorrelation.tex

\documentclass[aps,nofootinbib,pra,notitlepage,twocolumn]{revtex4-1}

\usepackage{amsfonts,amsmath,amssymb,amsthm}
\usepackage{array,bm,color}
\usepackage{epsfig,graphicx,nomencl,revsymb4-1,upgreek,url}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{graphicx}
% \usepackage[justification=justified,font=small]{caption}
\graphicspath{{./figures/}}
\hypersetup{colorlinks=true, pdfauthor=Kevin C. Young, pdftitle=Decorrelating Errors}
\newcommand{\tr}{{\rm Tr\thinspace}}
\newcommand{\bra}[1]{\ensuremath{\left\langle{#1}\right\vert}}
\newcommand{\ket}[1]{\ensuremath{\left\vert{#1}\right\rangle}}
\newcommand{\braket}[2]{\left\langle #1 | #2 \right\rangle}
\newcommand{\ketbra}[2]{\left| #1 \right\rangle\!\!\!\,\left\langle #2 \right|}
\newcommand{\abs}[1]{\left\vert #1 \right\vert}
\newcommand{\expect}[1]{\ensuremath{\left\langle{#1}\right\rangle}}
\newcommand{\timeorder}{\ensuremath{\underset{\leftarrow}{\mathcal{T}}}}
\newcommand{\ident}{{\mathbb1}}
\newcommand{\order}[1]{\mathcal{O}\left( #1 \right)}
\newcommand{\diag}[1]{\mathrm{diag}\{#1\}}
\newcommand{\trans}[1]{#1^\mathsf{T}}
\newcommand{\T}{\mathsf{T}}

\newcommand{\erf}[1]{Eq.~(\ref{#1})}
\newcommand{\needcite}{{\color{blue}\textsuperscript{[citation needed]}}}
\newcommand{\note}[1]{{\color{red}[#1]}}
\newcommand{\kcy}[1]{{\color{red}[#1]_{\rm{KCY}}}}
\newcommand{\amp}[1]{{\color{red}[#1]_{\rm{AMP}}}}
\setlength{\jot}{10pt}
%-------------Header begins here----------------------------------------
\begin{document}
% \tableofcontents
\title{Decorrelating Errors in Quantum Gates by Random Gate Synthesis - A BOCS of Control Solutions}

@@

Steady progress has been made in the theory of quantum error correction, proving higher thresholds for increasingly general models of noise \cite{Aharonov2006, 1609.00510, https://doi.org/10.7907/z96m34sc, Kubica2018, Wang2003, Campbell2017}. These results show that quantum computation is feasible in principle, however recent NISQ \cite{Preskill2018} devices have noise that is not only often above thresholds, but that also violates fundamental assumptions made by the models used in these results, such as Markovianity \cite{Kitaev1997} and independence of errors\cite{Knill1998}. With these assumptions violated, many properties of system performance and correctness can no longer be guaranteed.

To make matters worse, the ways in which a quantum computer can fail to meet these assumptions are manifold\cite{Kelly2018, BlumeKohout2017, Klimov2018}, and even when the noise is Markovian, many models and existing work make strong simplifying assumptions about the structure of the noise. As an example, Pauli channels are often used to model systems due to their classical simulability\cite{quant-ph/9807006}, even in the absence of physical motivation\cite{Aliferis2007, Knill2005, Wang2011, DuclosCianci2010, Wootton2012, Bombin2012, Puzzuoli2014, aliferis2008accuracy}. Because of this, a common approach is to attempt to use those thresholds to give a loose lower bound for thresholds with lower noise\cite{Puzzuoli2014}. This approach is correct and rigorous, but produces overly-pessimistic bounds on device performance.

Relatedly, many characterization routines assume that the noise is well-behaved so that they can make non-trivial assertions about performance. If these assumptions are violated, the utility of many of these algorithms is drastically reduced. For example, randomized benchmarking and tomography will report incorrect answers without any syndrome in the case of non-Markovian noise \cite{Merkel2013}. Slightly more helpful, gate-set tomography will report that the gateset failed to be Markovian, however figuring out how to bring it back to the space of the Markovianity can be challenging.

Many authors have approached these problems at varying levels of abstraction, each with their own benefits and short-comings. At the level of unitary synthesis, authors have considered randomizing over errors that might arise from gate approximations like Solovay-Kitaev \cite{Campbell2017, 1612.01011}. While this affords the ability to prove theorems that guarantee quadratic improvements to performance, they do not offer routines which are accessible and implementable with current coherence times. At the level of compilation, Frame Randomization\cite{Wallman2016, Ware2018}can be used to average over noise by \textit{twirling} error over an appropriately selected unitary 2-design.\cite{roy2009unitary} This has the advantage of being implementable\cite{Ware2018}, but it can require intensive use of classical resources, such as precompilation and generation of many waveform variants, and becomes more difficult when the gates being used are not in the Clifford group, as is the case in various quantum computing architectures. The general theme of these results is that of randomized decoupling\cite{Viola2005, Viola}. By introducing classical randomness into quantum computations, both coherent and non-Markovian can be transformed to incoherent, independent, Markovian noise, to varying degrees. In this paper, we explore a different method of solving these problems. We propose to inject additional decorrelating randomness into the system during physical gate synthesis through the use of \emph{balanced optimal control solutions} (BOCSs).

@@

\section{Theory}
 BOCSs are families of control solutions, $c_i(t)$, where each member of the family approximates the same target gate with slightly different errors, for any given instance a noise Hamiltonian. That is, for a target unitary operator $U_T$, we seek a family of approximating $U_i = \mathcal{T}e^{-i\int c_i(t)H(t, \delta)}$ to the target gate, such that the family of unitary approximations is \emph{balanced}. A balanced family is one which satisfies, for some small $\alpha$,
\begin{equation}\label{eq:1}
  \frac{1}{N}\sum_{i=1}^N \omega_i U_i \rho U_i^\dagger = DPN[\alpha]\left(U_T \rho U_T^\dagger \right)
\end{equation}

where $DPN[\alpha](\rho)$ is a \textit{generalized} depolarizing noise channel with strength $\alpha$. (For the rest of this paper, we will refer to them as just depolarizing channels.) Such a channel is defined as:
\begin{equation}\label{eq:2}
  DPN[\alpha](\rho) \rightarrow (1-\alpha)\rho + \alpha\sum p_i \sigma_i\rho\sigma_i
\end{equation}
with $p_i$ summing to one. This means that a for a family of controls that are a BOCS, there exists a weighting $p_i$, such that when the unitary approximations are averaged over with that weighting, they implement the target unitary followed by a depolarizing channel.

Thus, one property of BOCSs is that if each approximation of the target gate is correct to within $\epsilon$ in operator norm and $\alpha\leq\epsilon^2$, then the diamond distance of the BOCS to $U_T$ is also no larger than $\epsilon^2$. This follows from the mixing lemma proven by Campbell and Hastings in \cite{Campbell2017} and \cite{1612.01011}, and tells us that we can lower bound the worst case performance of a BOCS by the square of the error on each member of the BOCS, and linearly in the error of the weighted average. Because a set of parameters always exist that result in quadratically smaller error, it is then clear that in priniple BOCSs can always quadratically decrease the error in our approximation (in diamond norm). Moreover, in the setting where our system's performance is non-Markovian, e.g. drift, optimizing our BOCSs over a range of control parameters can reduce the non-Markovianity of the noise. As the name suggests, the task of generating BOCSs in this paper will fall to optimal control.

@@

\section{Optimal Control Problems}\label{ocp}
\subsection{Random Gate Synthesis}
Generating BOCSs can be done in a variety of ways, using any  of the many available quantum optimal control techniques \cite{Khaneja2005, Caneva2011, Machnes2018}. For our numerics, we chose to use the GRAPE algorithm to generate candidate pulseshapes to approximate the target gate. First described in \cite{Khaneja2005}, the GRAPE (GRadient Ascent Pulse Engineering) algorithm is a technique for finding piecewise constant control sequences that approximate a desired unitary, $U_T$. Defining our uncontrolled Hamiltonian as $H_0$, our control Hamiltonians as $H_{i\neq 0}$, and our \textit{control matrix} $c_{ij}$ as containing control amplitude associated with the $i^{th}$ time step and the $j^{th}$ hamiltonian, we can write our approximate unitary at any timestep as
\begin{equation}\label{eq:3}
  U_i = \exp\{-i\Delta t(H_0 + \sum_{j=1}^{n}c_{ij}H_{j}\}
\end{equation}
Then, to measure the simularity of our approxiate unitary $U_n$ to our target unitary $U_T$, we can define a cost function $J(U) = Tr\{U_T^{\dagger}U_n\}$.

To optimize this cost function we can perform the following standard update loop for some threshold value $\varepsilon > 0$ and step size $\delta > 0$:
\begin{algorithm}[H]
\floatname{algorithm}
  \caption{\textsc{\textbf{Gradient Ascent}}}
  \begin{algorithmic}
    \While{$J(U_n) < (1-\varepsilon$)}
    \State $c_{ij} \rightarrow c_{ij} + \delta\frac{\partial J(U)}{\partial c_{ij}}$
    \For{$1 \leq i \leq n$}
    \State $U_i \rightarrow \exp\{-i\Delta t(H_0 + \sum_{i=0}^{n}c_{ij}H_j)\}$
    \EndFor
    \State $U \rightarrow \prod_{i=1}^nU_i$
    \EndWhile
  \end{algorithmic}
\end{algorithm}

@@

In our numerical results in Section \ref{1Q Gates} and \ref{2Q Gates}, we consider the controls to have quasi-static Gaussian distributed errors, so to generate controls that are robust drift in these controls we modify our gradient to instead be:
\begin{align}\label{quadrature}
\frac{\partial \tilde J(U)}{\partial u_{ij}} =
\int p(\vec{\delta})\frac{\partial J(U(\vec{\delta}))}{\partial u_{ij}} d\vec{\delta}
\end{align}
with $p(\vec{\delta})$ Gaussian distributed, as has been done in previous works \cite{Goerz2014} to ensure that the optimal control results are robust over a wide range of errors. To make this averaging tractable, we approximate this integral using Gaussian quadrature, approximating the cost functions as a low order polynomial.

Concretely, we consider a Hamiltonian of the following form:
\begin{equation}\label{eq:2}
  H(t) = \delta_0H_0 + \sum_{i=1}^n (1 + \delta_i)c_i(t)H_i
\end{equation}
for control Hamiltonians $H_i$, free evolution Hamiltonian $H_0$ and random variables $\delta_i$, that model some small uncertainty in parameters in the Hamiltonian. Such a model might describe a superconducting qubit quantum processor where control amplitudes for the RF pulses vary over time, or a trapped ion quantum computer where the intensity, frequency, or phase of the laser might drift over time\cite{Kelly2018, Lekitsch2017}. Correlations between different $\delta_i$ might arise, for instance, if two of the controls have the same noise source, e.g. $RX$ and $RY$ gates in superconducting qubit architectures might use the same AWG and pulse envelope, and suffer from the same diurnal temperature drift of control electronics.

\subsection{BOCS Approximation}
After using GRAPE or another optimal control routine to synthesize a collection of controls, we must find the weights $w_i$ such that the collection of controls form a BOCS as described in Equation \ref{eq:1}. To do this, for each control $U_i$ we find the unitary error channel $\mathcal{E}_i$ such that $\mathcal{E}_iU_i=U_T$, where $U_T$ is the target gate. The we see that for any stochastic application of these channels, the resulting map is given by:
\begin{align}
 \frac{1}{N} \sum^N_{i=1} w_i \mathcal{E}_i^{\dagger} (U_T\rho U_T^{\dagger}) \mathcal{E}_i
\end{align} If we consider the Pauli-Liouville representation\cite{Kimmel2014} of this error channel, the diagonal terms are the \textit{stochastic} terms that arise from classical uncertainty, while the off-diagonal terms may more generally arise from \textit{coherent} errors.
Thus to approximate a depolarizing channel we define our optimal control problem to be the following, which minimizes the off-diagonal terms:
\begin{equation}\label{eq:minimization}
\setlength{\jot}{0pt}
  \begin{split}
    &\underset{w_0, ..., w_N}{\textbf{minimize}} \{\sum_{\substack{i,j \\ i\neq j}}^N|\sigma_i\Lambda(\sigma_j)|^2\}\\
    &\textbf{where}\ \Lambda(\sigma_j) := \sum^N_{i=1}w_i\mathcal{E}_i^{\dagger}\sigma_j\mathcal{E}_i\\
    &\textbf{subject to} \sum_{i=1}^Nw_i = 1
  \end{split}
\end{equation}

@@

Previous authors have considered minimizing the diamond distance to the nearest Pauli or Clifford Channel \cite{Magesan2013}, and while this gives a good theoretical framework, it requires the more computationally challenging task of optimizing over the diamond norm, and does not constrain the resulting channel to be decomposable into a given family of controls. Our routine, on the other hand, optimizes over an easy to compute sum, and produces a channel defined in terms of given collection of pulseshapes.

@@

% \begin{figure*}
% \centering
% \begin{subfigure}[t]{.5\linewidth}
% \includegraphics[width=\textwidth]{control_dpn_all0.png}
% \caption{This is one example of a 1D slice varying over one control. There are five of these in total, but only one is interesting.}
% \end{subfigure}%
% ~
% \begin{subfigure}[t]{.5\linewidth}
% \includegraphics[width=\textwidth]{control_dpn_all2.png}
% \caption{This is one example of a 1D slice varying over one control. There are five of these in total, but only one is interesting.}
% \end{subfigure}
%   \label{fig:2qnum}
% \end{figure*}
\section{Numerical Results}\label{numerical}
In the following two subsections, we present numerical results of our routine, first for a one qubit example, and then for a two-qubit example. The code and BOCSs generated for both examples is available online at \cite{decorrelating_errors}.
\subsection{1Q Gates}\label{1Q Gates}
 For the one-qubit case, we consider generating BOCSs for $RX(\frac{\pi}{2})$ and $RY(\frac{\pi}{2})$, that together with $RZ(\theta$) rotations are universal for one-qubit computation. (In particular, choosing just one of these two gates would be sufficient.) Our control Hamiltonian is given as:
\begin{equation}\label{eq:1Qham}
  H = \epsilon\sigma_z + (1 + \delta)(c_x(t)\sigma_x + c_y(t)\sigma_y)
\end{equation}
where $\epsilon, \delta \sim \mathcal{N}(0, .001)$ We assume that the errors on $\sigma_x$ and $\sigma_y$ are perfecly correlated, as mentioned in Section \ref{ocp}. In our simulation we chose an total evolution time of $T=\pi$, a number of steps $N=100$, and a threshold infidelity of $1E-3$ for generating the control solutions with GRAPE.

@@

\subsection{2Q Gates}\label{2Q Gates}
 For a two-qubit example, we again consider the single qubit gates $RX(\frac{\pi}{2})$ and $RY(\frac{\pi}{2})$ on both qubits, along with single qubit $RZ(\theta)$ rotations. The entangling operation we chose is $ZZ(\frac{\pi}{2})$, which together with the single-qubit operations is universal for quantum computation. Our control Hamiltonian is given as:
\begin{equation} \label{eq:2Qham}
\begin{split}
H = &\sum_{j=1}^2(\epsilon_j\sigma_z^j + (1 + \delta_j)(c_x^jx(t)\sigma_x^j + c_y^j(t)\sigma_y^j)) \\
&+ \Delta c(t) \exp{(-i\frac{\sigma_z^1\otimes\sigma_z^2}{4})}
\end{split}
\end{equation}
We again consider $\epsilon_j, \delta_j, \delta \sim .001$. In this simulation we again had a threshold infidelity of $1E-3$, but we increased the total evolution time to $T=4\pi$, and increased the number of steps to $N=400$ so that the size of each time step was the same as in the one qubit example, however the total evolution time was greater to allow GRAPE more opportunities to find non-trivial pulseshapes.

%0.0507,  0.0307, -0.0493, -0.0693, with total area under the curve given by 0.7853981633974484, 
\section{Experimental Results}\label{experimental}
Here we present experimental results from implementing our routine on a fixed-frequency superconducting transmon qubit. In particular we used qubit 8 on the Rigetti 19Q-Acorn chip, whose characterization can be found in \cite{1712.05771}. To implement a BOCS on this qubit, four incorrectly calibrated approximately Gaussian pulses were produced by scaling the pulseshape for an approximately calibrated 50$\mu$s $X_{\frac{\pi}{2}}$ pulse by $106.4\%$,  $103.9\%$, $93.7\%$ and $91.2\%$ pulse.

Using Equation \ref{eq:minimization}, we then generated the optimal (Is this convex?) weights $\omega_i$, and ran a randomized benchmarking experiment for each over- and under-calibrated pulse, the calibrated pulse, and the stochastic channel given by drawing from the members of the BOCS with $\omega_i$ on each application of the gate. We used $N=1000$ shots per experiment and $K=10$ sequences per sequence length, out to length $L=64$\cite{Magesan2011}. In each case, our Clifford operations were decomposed into RX($\frac{\pi}{2})$) and RY($\frac{\pi}{2})$) pulses. In our implementation, these gates are implemented using the same pulse envelope definitions with the same control electronics, phase shifted by $\frac{\pi}{2}$ radians, and are therefore subject to the same miscalibration errors.
 It was shown in \cite{Ball2016} that for particular non-Markovian (\textit{quasi-static}) error models, noise will manifest as gamma distributed points for each sequence length. On the other hand, Markovian noise, such as depolarizing noise, will result in Gaussian distributed fidelity estimates for each randomized benchmarking sequence length. As can be seen in the results shown in Figure \ref{fig:rb} for sequence lengths $L=64$, we see that the coherently miscalibrated controls having long tails, consistent with gamma distributed random variables, while the calibrated andn randomized implementations both have much shorter tails, consistent with gaussian distributed random variables.

>>>>

\documentclass[aps,nofootinbib,pra,notitlepage,twocolumn]{revtex4-1}
\usepackage{amsfonts,amsmath,amssymb,amsthm}
\usepackage{array,bm,color}
\usepackage{epsfig,graphicx,nomencl,revsymb4-1,upgreek,url}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{graphicx}
% \usepackage[justification=justified,font=small]{caption}
\graphicspath{{./figures/}}
\hypersetup{colorlinks=true, pdfauthor=Kevin C. Young, pdftitle=Decorrelating Errors}
\newcommand{\tr}{{\rm Tr\thinspace}}
\newcommand{\bra}[1]{\ensuremath{\left\langle{#1}\right\vert}}
\newcommand{\ket}[1]{\ensuremath{\left\vert{#1}\right\rangle}}
\newcommand{\braket}[2]{\left\langle #1 | #2 \right\rangle}
\newcommand{\ketbra}[2]{\left| #1 \right\rangle\!\!\!\,\left\langle #2 \right|}
\newcommand{\abs}[1]{\left\vert #1 \right\vert}
\newcommand{\expect}[1]{\ensuremath{\left\langle{#1}\right\rangle}}
\newcommand{\timeorder}{\ensuremath{\underset{\leftarrow}{\mathcal{T}}}}
\newcommand{\ident}{{\mathbb1}}
\newcommand{\order}[1]{\mathcal{O}\left( #1 \right)}
\newcommand{\diag}[1]{\mathrm{diag}\{#1\}}
\newcommand{\trans}[1]{#1^\mathsf{T}}
\newcommand{\T}{\mathsf{T}}

\newcommand{\erf}[1]{Eq.~(\ref{#1})}
\newcommand{\needcite}{{\color{blue}\textsuperscript{[citation needed]}}}
\newcommand{\note}[1]{{\color{red}[#1]}}
\newcommand{\kcy}[1]{{\color{red}[#1]_{\rm{KCY}}}}
\newcommand{\amp}[1]{{\color{red}[#1]_{\rm{AMP}}}}
\setlength{\jot}{10pt}
%-------------Header begins here----------------------------------------
\begin{document}
% \tableofcontents
\title{Decorrelating Errors in Quantum Gates by Random Gate Synthesis}

@@

Steady progress has been made in the theory of quantum error correction, proving higher thresholds for increasingly general models of noise \cite{Aharonov2006, 1609.00510, https://doi.org/10.7907/z96m34sc, Kubica2018, Wang2003, Campbell2017}. These results show that quantum computation is feasible in principle, however recent NISQ \cite{Preskill2018} devices have noise that is not only often above known thresholds, but that also violates fundamental assumptions made by the models used in these results\cite{Kelly2018, BlumeKohout2017, Klimov2018}, such as Markovianity \cite{Kitaev1997} and independence of errors\cite{Knill1998}. With these assumptions violated, many properties of system performance and correctness can no longer be guaranteed.

Moreover, even when noise can be accurately modeled as Markovian, many existing works make further simplifying approximations about its structure to make problems tractable. For example, Pauli channels are often used to model systems due to their classical simulability\cite{quant-ph/9807006}, even in the absence of physical motivation\cite{Aliferis2007, Knill2005, Wang2011, DuclosCianci2010, Wootton2012, Bombin2012, Puzzuoli2014, aliferis2008accuracy}. One approach, then, would be to use these thresholds to give a loose lower bound for thresholds in other systems\cite{Puzzuoli2014}. Such an approach is correct and rigorous, but will produce overly-pessimistic bounds that may not be reflective of numerically demonstrated performnace.

Many characterization routines also make assumptions about the form of noise, and if these assumptions are violated the utility and correctness of these routines may be reduced. For example, randomized benchmarking and tomography will report incorrect answers without any syndrome in the case of non-Markovian noise \cite{Merkel2013}. Slightly more helpful, gate-set tomography will report that a system fails to be Markovian, without saying in what way..

Other authors have approached these problems at varying levels of abstraction, each with their own benefits and short-comings. At the level of unitary synthesis, one can consider randomizing over errors that might arise from gate approximations like the Solovay-Kitaev decomposition \cite{Campbell2017, 1612.01011}. While this affords the ability to prove theorems that guarantee quadratic improvements to performance, they do not offer routines which are accessible and implementable with current coherence times. In addition, present quantum computing work focusses primarily on gate synthesis, rather than unitary synthesis.

At the level of compilation, Frame Randomization\cite{Wallman2016, Ware2018} can be used to average over noise by twirling error over an appropriately selected unitary 2-design.\cite{roy2009unitary} This has the advantage of being realizable in existing systems\cite{Ware2018}, but it can require intensive use of classical resources, such as precompilation and generation of many waveform variants, and becomes more difficult when the gates being used are not in the Clifford group, as is the case in various quantum computing architectures.

The general theme of these results is randomized decoupling\cite{Viola2005, Viola} -- by introducing classical randomness into quantum computations, both coherent and non-Markovian noise can be transformed to incoherent, independent, and Markovian noise. In this paper, we explore a different method of solving these problems. We propose to inject additional decorrelating randomness into the system during physical gate synthesis through the use of \emph{balanced optimal control solutions} (BOCSs). The task of generating BOCSs will fall to optimal control.

@@

\section{Theory}
 BOCSs are families of control solutions, $c_i(t)$, where each member of the family approximates the same target gate with slightly different errors, for any given instance a noise Hamiltonian. That is, for a target gate $U_T$, we seek a family of approximating gates $U_i = \mathcal{T}e^{-i\int c_i(t)H(t, \delta)}$ such that the family of unitary approximations is \emph{balanced}. A family of controls is balanced if there is a weighting $\omega_i$ of its members such that produces a \textit{balanced channel}, satisfying for some small $\alpha$,
\begin{equation}\label{eq:1}
  \frac{1}{N}\sum_{i=1}^N \omega_i U_i \rho U_i^\dagger = DPN[\alpha]\left(U_T \rho U_T^\dagger \right)
\end{equation}

where $\omega_i > 0$ and $DPN[\alpha](\rho)$ is a depolarizing noise channel, with strength $\alpha$. Such a channel is defined as:
\begin{equation}\label{eq:2}
  DPN[\alpha](\rho) \rightarrow (1-\alpha)\rho + \alpha\sum p_i \sigma_i\rho\sigma_i
\end{equation}
with $p_i$ summing to one.

Thus, following from the mixing lemma proven by Campbell and Hastings in \cite{Campbell2017} and \cite{1612.01011}, one property of BOCSs is that if each approximation of the target gate is correct to within $\epsilon$ in operator norm and $\alpha\leq\epsilon^2$, then the diamond distance of the BOCS to $U_T$ is also no larger than $\epsilon^2$. Because any sufficiently populated BOCS will have a set of controls that result in error of order $\mathcal{O}(\epsilon^2)$ on their balanced channel, it is then clear that in priniple BOCSs can always quadratically decrease the error in our approximation (in diamond norm). Moreover, in the setting where our system's performance is non-Markovian, e.g. drift, optimizing our BOCSs \textit{robustly} over a range of control parameters can reduce the non-Markovianity of the noise.

@@

\section{Optimal Control Problems}\label{ocp}
\subsection{Random Gate Synthesis}
Generating BOCSs can be done in a variety of ways, using any  of the many available quantum optimal control techniques \cite{Khaneja2005, Caneva2011, Machnes2018}. For our numerics, we chose to use the GRAPE algorithm to generate candidate pulseshapes to approximate the target gate. First described in \cite{Khaneja2005}, the GRAPE (GRadient Ascent Pulse Engineering) algorithm is a technique for finding piecewise constant control sequences that approximate a desired unitary, $U_T$. Defining our uncontrolled Hamiltonian as $H_0$, our control Hamiltonians as $H_{i\neq 0}$, and our \textit{control matrix} $c_{ij}$ as containing control amplitude associated with the $i^{th}$ time step and the $j^{th}$ hamiltonian, we can write our approximate unitary at any timestep as
\begin{equation}\label{eq:3}
  U_i = \exp\{-i\Delta t(H_0 + \sum_{j=1}^{n}c_{ij}H_{j}\}
\end{equation}
Then, to measure the simularity of our approxiate unitary $U$ to our target unitary $U_T$, we can define a cost function $J(U) = Tr\{U_T^{\dagger}U\}$.

To optimize this cost function we can perform the following standard update loop for some threshold value $\varepsilon > 0$ and step size $\delta > 0$:
\begin{algorithm}[H]
\floatname{algorithm}
  \caption{\textsc{\textbf{Gradient Ascent}}}
  \begin{algorithmic}
    \While{$J(U) < (1-\varepsilon$)}
    \State $c_{ij} \rightarrow c_{ij} + \delta\frac{\partial J(U)}{\partial c_{ij}}$
    \For{$1 \leq i \leq n$}
    \State $U_i \rightarrow \exp\{-i\Delta t(H_0 + \sum_{i=0}^{n}c_{ij}H_j)\}$
    \EndFor
    \State $U \rightarrow \prod_{i=1}^nU_i$
    \EndWhile
  \end{algorithmic}
\end{algorithm}

@@

In our numerical results in Section \ref{1Q Gates} and \ref{2Q Gates}, we consider the controls to have quasi-static\cite{Ball2016}(non-Markovian) Gaussian distributed errors, so to generate controls that are robust to this error we modify our gradient to instead be:
\begin{align}\label{quadrature}
\frac{\partial \tilde J(U)}{\partial u_{ij}} =
\int p(\vec{\delta})\frac{\partial J(U(\vec{\delta}))}{\partial u_{ij}} d\vec{\delta}
\end{align}
with $p(\vec{\delta})$ Gaussian distributed, as has been done in previous works \cite{Goerz2014} to ensure that the optimal control results are robust over a wide range of errors. To make this averaging tractable, we approximate this integral using Gaussian quadrature, approximating the cost functions as a low order polynomial\cite{abramowiz1972handbook}.

Concretely, we consider a Hamiltonian of the following form:
\begin{equation}\label{eq:2}
  H(t) = \delta_0H_0 + \sum_{i=1}^n (1 + \delta_i)c_i(t)H_i
\end{equation}
for control Hamiltonians $H_i$, free evolution Hamiltonian $H_0$ and random variables $\delta_i$, that model some small uncertainty in parameters in the Hamiltonian. Such a model might describe a superconducting qubit quantum processor where control amplitudes for the RF pulses or qubit transition frequencies vary over time, or a trapped ion quantum computer where the intensity, frequency, or phase of the laser might drift\cite{Kelly2018, Klimov2018, Lekitsch2017}. Correlations between different $\delta_i$ might arise, for instance, if two of the controls have the same noise source, e.g. $RX$ and $RY$ gates in superconducting qubit architectures might use the same AWG and pulse envelope, and all of the lab equipment may suffer from the same temperature drift of control electronics.

\subsection{BOCS Approximation}
After using GRAPE or another optimal control routine to synthesize a collection of controls, we must find the weights $\omega_i$ such that the collection of controls form a BOCS as described in Equation \ref{eq:1}. To do this, for each control $U_i$ we find the unitary error channel $\mathcal{E}_i$ such that $\mathcal{E}_iU_i=U_T$, where $U_T$ is the target gate. The we see that for any stochastic application of these channels, the resulting map is given by:
\begin{align}
 \frac{1}{N} \sum^N_{i=1} w_i \mathcal{E}_i^{\dagger} (U_T\rho U_T^{\dagger}) \mathcal{E}_i
\end{align} If we consider the Pauli-Liouville representation\cite{Kimmel2014} of this error channel, the diagonal terms are the \textit{stochastic} terms that arise from classical uncertainty, while the off-diagonal terms may more generally arise from \textit{coherent} errors\cite{Ware2018}.
Thus to approximate a depolarizing channel we define our optimal control problem to be the following, which minimizes the off-diagonal terms:
\begin{equation}\label{eq:minimization}
\setlength{\jot}{0pt}
  \begin{split}
    &\underset{w_0, ..., w_N}{\textbf{minimize}} \{\sum_{\substack{i,j \\ i\neq j}}^N|\sigma_i\Lambda(\sigma_j)|^2\}\\
    &\textbf{where}\ \Lambda(\sigma_j) := \sum^N_{i=1}w_i\mathcal{E}_i^{\dagger}\sigma_j\mathcal{E}_i\\
    &\textbf{subject to} \sum_{i=1}^Nw_i = 1
  \end{split}
\end{equation}

@@

Previous authors have considered minimizing the diamond distance to the nearest Pauli or Clifford Channel \cite{Magesan2013}, and while this gives a good theoretical framework, it requires the more computationally challenging task of optimizing over the diamond norm, and does not constrain the resulting channel to be decomposable into a given family of controls. Our routine, on the other hand, optimizes over an easy to compute sum, and produces a channel defined explicitly in terms of given family of controls.

@@

% \begin{figure*}
% \centering
% \begin{subfigure}[t]{.5\linewidth}
% \includegraphics[width=\textwidth]{control_dpn_all0.png}
% \caption{This is one example of a 1D slice varying over one control. There are five of these in total, but only one is interesting.}
% \end{subfigure}%
% ~
% \begin{subfigure}[t]{.5\linewidth}
% \includegraphics[width=\textwidth]{control_dpn_all2.png}
% \caption{This is one example of a 1D slice varying over one control. There are five of these in total, but only one is interesting.}
% \end{subfigure}
%   \label{fig:2qnum}
% \end{figure*}

\section{Numerical Results}\label{numerical}
In the following two subsections, we present numerical results of our routine, first for a one qubit example, and then for a two-qubit example. The code and BOCSs generated for both examples are available online at \cite{decorrelating_errors}.
\subsection{1Q Gates}\label{1Q Gates}
 For the one-qubit case, we generate BOCSs for $RX(\frac{\pi}{2})$ and $RY(\frac{\pi}{2})$, that together with $RZ(\theta$) rotations are universal for one-qubit computation. (In particular, choosing just one of these two gates would be sufficient.) Our control Hamiltonian is given as:
\begin{equation}\label{eq:1Qham}
  H = \epsilon\sigma_z + (1 + \delta)(c_x(t)\sigma_x + c_y(t)\sigma_y)
\end{equation}
where $\epsilon, \delta \sim \mathcal{N}(0, .001)$ We assume that the errors on $\sigma_x$ and $\sigma_y$ are perfecly correlated, as mentioned in Section \ref{ocp}. In our simulation we chose an total evolution time of $T=\pi$, a number of steps $N=100$, and a threshold infidelity of $1E-3$ for generating the control solutions with GRAPE.

@@

\subsection{2Q Gates}\label{2Q Gates}
 For a two-qubit example, we again consider the single qubit gates $RX(\frac{\pi}{2})$ and $RY(\frac{\pi}{2})$ on both qubits, along with single qubit $RZ(\theta)$ rotations. The entangling operation we chose is $ZZ(\frac{\pi}{2})$, which together with the single-qubit operations are universal for quantum computation. Our control Hamiltonian is given as:
\begin{equation} \label{eq:2Qham}
\begin{split}
H = &\sum_{j=1}^2(\epsilon_j\sigma_z^j + (1 + \delta_j)(c_x^jx(t)\sigma_x^j + c_y^j(t)\sigma_y^j)) \\
&+ (1+\delta)c(t) \exp{(-i\frac{\sigma_z^1\otimes\sigma_z^2}{4})}
\end{split}
\end{equation}
We again consider $\epsilon_j, \delta_j, \delta \sim .001$. In this simulation we again used a threshold infidelity of $1E-3$, but we increased the total evolution time to $T=4\pi$, and increased the number of steps to $N=400$ so that the size of each time step was the same as in the one qubit example, however the total evolution time was greater to allow GRAPE more opportunities to find non-trivial pulseshapes.

%0.0507,  0.0307, -0.0493, -0.0693, with total area under the curve given by 0.7853981633974484, 
\section{Experimental Results}\label{experimental}
Here we present experimental results from implementing our routine on a fixed-frequency superconducting transmon qubit. In particular we used qubit 8 on the Rigetti 19Q-Acorn chip, whose characterization can be found in \cite{1712.05771}. To implement a BOCS on this qubit, four incorrectly calibrated approximately Gaussian pulses were produced by scaling the pulseshape for a calibrated 50$\mu$s $RX(\frac{\pi}{2})$ pulse by $106.4\%$,  $103.9\%$, $93.7\%$ and $91.2\%$.

Using Equation \ref{eq:minimization}, we then generated the optimal (IS THIS CONVEX?) weights $\omega_i$. To benchmark the quality of the new balanced channel, we then performed six randomized benchmarking experiments: one for each over- and under-calibrated pulse, one for the calibrated pulse, and one for the balanced channel. We used $N=1000$ shots per experiment and $K=10$ sequences per sequence length, for sequence lengths up to $L=64$\cite{Magesan2011}. In each case, our Clifford operations were decomposed into RX($\frac{\pi}{2})$) and RY($\frac{\pi}{2})$) pulses. In our implementation, these gates are implemented using the same pulse envelope definitions and control electronics, phase shifted by $\frac{\pi}{2}$ radians, and are therefore subject to identical miscalibration errors. The results are shown in Figure \ref{fig:rb} for sequence lengths $L=64$. In this experiment, fitting the randomized benchmarking data with scipy's curve\_fit using the Trust Region Reflective method reports one-qubit gate fidelities of $99.4\%$ for the calibrated pulse, $98.8\%$ for Pulse1, $99.4\%$ for Pulse2, $98.9\%$ for Pulse3, $98.6\%$ for Pulse4, and $99.0\%$ for the Randomized pulse. As one can see from looking at Figure \ref{fig:rb}, despite all controls performing comparably to the Calibrated control, only the Randomized control has similarly tight error bars. In the cases of each miscalibrated pulse, the error bars extend significantly, in often asymmetric ways, making usage of chi-square statistics, and errors on the fit less straightforward.

This is consistent with the result shown in \cite{Ball2016} that for particular non-Markovian error models, noise will manifest as gamma distributed points for each sequence length. On the other hand, Markovian noise, such as depolarizing noise, will result in Gaussian distributed fidelity estimates for each randomized benchmarking sequence length. We see that the coherently miscalibrated controls have long tails, consistent with gamma distributed random variables, while the calibrated and randomized implementations both have much shorter tails, consistent with Gaussian distributed random variables.
